Why classical mechanics fails

### 2023-05-19 Articles/2021-WhyClassicalMechanicsFails/WhyClassicalMechanicsFails.tex
Fixed typos

Before we explore more what happens across the system-environment boundary, we note that quantum mechanics does suffer from the same problem. The entropy of any pure state in quantum mechanics is zero, not minus infinity. The entropy of any mixed quantum state is greater than zero. Therefore there are no quantum states with negative entropy. In this light, we can picture the classical description becoming more and more untenable as entropy decreases. That is, in the same way that we understand non-relativistic mechanics as the limit of relativistic mechanics when velocities are much smaller than $c$, we should understand classical mechanics as the limit of quantum mechanics for states whose entropy is much greater than $0$.

@@

The problem with classical mechanics is not so much that it assumes perfect isolation, but that it assumes it at all scales. Classical fluids and continuous materials allow us to study smaller and smaller parts. Classical fields allow for wave packets of ever decreasing energy. The theory implicitly assumes all these objects can be prepared, manipulated and detected independently enough that their state can be perfectly defined. This is the inherent problem with classical mechanics: perfect isolation at all scales. In the correct circumstances, perfect isolation is an internally consistent assumption. In the correct circumstances, recursive subdivision into smaller parts is an internally consistent assumption. It is the joint assumption that the system can be recursively subdivided into perfectly isolated parts that is physically untenable.\footnote{Note that the mathematical model is still perfectly consistent. It is the physical model that is not. That is why one should never confuse mathematical consistency with physical consistency.} As we consider smaller and smaller scales, study smaller and smaller objects at an ever increasing level of precision, we are inevitably bound to move away from the regime where perfect isolation can be assumed.

@@

We have shown that classical mechanics lies on the implicit assumption of perfect isolation at all scales: every fragment of matter, every minuscule wave packet, can be produced, manipulated and measured with infinitesimal precision. This is not only untenable, but it is also in conflict with thermodynamics. The existence of an unavoidable minimal interaction with the environment, with the uncertainty associated with it, is ultimately what makes the classical ideas fail.

>>>>

Before we explore more what happens across the system-environment boundary, we note that quantum mechanics does not suffer from the same problem. The entropy of any pure state in quantum mechanics is zero, not minus infinity. The entropy of any mixed quantum state is greater than zero. Therefore there are no quantum states with negative entropy. In this light, we can picture the classical description becoming more and more untenable as entropy decreases. That is, in the same way that we understand non-relativistic mechanics as the limit of relativistic mechanics when velocities are much smaller than $c$, we should understand classical mechanics as the limit of quantum mechanics for states whose entropy is much greater than $0$.

@@

The problem with classical mechanics is not so much that it assumes perfect isolation, but that it assumes it at all scales. Classical fluids and continuous materials allow us to study smaller and smaller parts. Classical fields allow for wave packets of ever decreasing energy. The theory implicitly assumes all these objects can be prepared, manipulated and detected independently enough that their state can be perfectly defined. This is the inherent problem with classical mechanics: perfect isolation at all scales. In the correct circumstances, perfect isolation is a reasonable assumption. In the correct circumstances, recursive subdivision into smaller parts is a reasonable assumption. It is the joint assumption that the system can be recursively subdivided into perfectly isolated parts that is physically untenable.\footnote{Note that the mathematical model is still perfectly consistent. It is the physical model that is not. That is why one should never confuse mathematical consistency with physical consistency.} As we consider smaller and smaller scales, study smaller and smaller objects at an ever increasing level of precision, we are inevitably bound to move away from the regime where perfect isolation can be assumed.

@@

We have shown that classical mechanics depends on the implicit assumption of perfect isolation at all scales: every fragment of matter, every minuscule wave packet, can be produced, manipulated and measured with infinitesimal precision. This is not only untenable, but it is also in conflict with thermodynamics. The existence of an unavoidable minimal interaction with the environment, with the uncertainty associated with it, is ultimately what makes the classical ideas fail.
